### 2017-10-05 SP_data_cleaning.Rmd
Add spanish data and initial exploration

---
title: "SP"
author: "I. Bartomeus"
date: "05/10/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Notas:

- 'Localidades?': Info per species. Recorded only at the first record of each species. Refers to geograpfic known records. Usefull for abejassilvestres.es ?
- 'obrera' needs to be merged with 'sex'
- ID and 'cuaderno' need to remove duplicates, which are not in cuaderno!!
- IF several IT's Oscar Aguade duplicated the netry as many times as need. Those sp are in red in original excel.

Old data:

```{r}
sp <- read.csv(file = "raw_data/Spain.csv", h = T)
head(sp)

```

New data:

- Not entered yet.