### 2018-02-07 SP_data_cleaning.Rmd
Start preparing Spain data

---
title: "SP"
author: "I. Bartomeus"
date: "05/10/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Notas:

- 'Localidades?': Info per species. Recorded only at the first record of each species. Refers to geograpfic known records. Usefull for abejassilvestres.es ?
- 'obrera' needs to be merged with 'sex'
- ID and 'cuaderno' need to remove duplicates, which are not in cuaderno!!
- IF several IT's Oscar Aguado duplicated the enetry as many times as need. Those sp are in red in original excel.

Old data:

```{r}
sp <- read.csv(file = "raw_data/Spain.csv", h = T)
head(sp)
head(sort(table(sp$LOCALIDAD), decreasing = TRUE))
write.csv(sort(table(sp$LOCALIDAD), decreasing = TRUE),
          "raw_data/localidades.csv")

sp$gen_sp <- paste(sp$G.NERO, sp$ESPECIE, sep = "_")
unique(sp$gen_sp) #721 species! #clean, but only for the subset of used data. The 7 localities.
head(sort(table(sp$gen_sp), decreasing = TRUE), 50)

#subset only the 7 localities.

#check sampling completness
library(vegan)
#need to fix first locality
comm <- table(sp2$LOCALIDAD, sp2$gen_sp)
temp <- specaccum(comm, "random")
summary(temp)
plot(temp)
rarecurve(comm)

```

New data:

```{r}
now <- read.csv(file = "raw_data/Spain_2016.csv", h = T)
head(now)
table(now$LOCALIDAD)
levels(now$LOCALIDAD) <- c("auditorio_Delibes",
                           "auditorio_Delibes",
                           "hornillos",
                           "hornillos",
                           "olmedo",
                           "sanbernardo",
                           "toro",
                           "vilalba",
                           "zamaduenas")
#get lat long
library(spatial)
#biogeomancer(country = "Spain", adm1 = "Castilla y Leon", adm2 = "Valladolid", 
#             locality = levels(now$LOCALIDAD), 
#             singleRecord=TRUE, progress="text")
#not working, but Ropen sci has alternative packages.
now$gen_sp <- paste(now$GÉNERO, now$ESPECIE, sep = "_")
sort(table(now$gen_sp))
#clean species
unique(now$gen_sp) #197!!! but still some duplicates

#check sampling completness
library(vegan)
comm <- table(now$LOCALIDAD, now$gen_sp)
temp <- specaccum(comm, "random")
summary(temp)
plot(temp)
rarecurve(comm) #uf... not really well sampled... but ~80 species per site it's quite good!
```

Data Gbif:
```{r}
gbif <- read.csv(file = "raw_data/data_gbif.csv", h = T)
head(gbif)
hist(gbif$year) #nice, bimodal!
sort(table(gbif$recordedBy), decreasing = TRUE)
#Engel, M, Hinojosa, I, Bennett, D, & Davis, S
#G.E. Bohart
#Torchio, Asensio, etc...
#Baker, D. & Baker, M.
# see
set <- which(gbif$recordedBy == "Engel, M, Hinojosa, I, Bennett, D, & Davis, S")
library(mapr)
colnames(gbif)[2:3] <- c("latitude", "longitude") 
map_leaflet(gbif[set,]) #interesting, 
# Engel collected in 2007 in the south! And taxonomy is bad!
# Baker in Estartit #del 62! 
# Bohart Madrid, islands, south... in 78's

#Overall, most pre 1980 collections around valladolid.
head(sort(table(gbif$species), decreasing = TRUE), 50)
unique(gbif$species) #443! BUT here lots of duplicates?
```

#Analysis

1) only for 7 localities, join now and then and paired t-test on rarefied richness per site

2) For common species we can do the same but with standardized frequency.