### 2018-03-26 ToDo.Rmd
cleaning repo

---
title: "Untitled"
author: "I. Bartomeus"
date: "26/03/2018"
output: html_document
---

To Do:

- check formating rules (words in abstract, main body, etc...)
- correct fig1 for country size?
- Species declines in % for different taxa... what to do?
- think on gbif plot...
- Spain.

#refs clarifications

@Ollerton2012 
@Winfree2011 - >